```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Atomic vector (", as.character(.version), ")", sep="")
```

## Overview

The `atomic_vector` format provides an on-disk representation of typed arrays, based on R's atomic vectors.

## Metadata

The top-level object in the `OBJECT` file should contain an `atomic_vector` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `atomic_vector` file format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain a `contents.h5` HDF5 file.
This file should contain an `atomic_vector` group with the following attributes:

- `type`, a scalar string attribute specifying the type of the vector.
  This should be one of `"integer"`, `"boolean"`, `"string"` or `"number"`.

The `atomic_vector` group should in turn contain `values`, a 1-dimensional dataset.
The HDF5 datatype of the dataset should match the `type` attribute:

- For `type = "integer"` or `"boolean"`, the datatype should be any integer type that can be exactly represented by a 32-bit signed integer.
  Check out the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for more details.
- For `type = "number"`, the datatype should be any integer/float type that can be exactly represented by a 64-bit float.
- For `type = "string"`, the datatype should be any string type that can be represented by a UTF-8 encoded string.

Missing values are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `values` dataset.
All values in the dataset equal to the placeholder should be treated as missing.
The placeholder datatype should be identical to that of `values`, except in the case of strings where any string datatype may be used.

For `type = "string"`, the `atomic_vector` group may optionally contain a `format` attribute.
This should be set to:

- `"none"`: no constraints on the contents of each string.
  This is the default behaviour if no `format` is present.
- `"date"`: strings should be dates following a `YYYY-MM-DD` format.
- `"date-time"`: strings should be Internet Date/Time values following the format described in [RFC3339](https://www.rfc-editor.org/rfc/rfc3339).

Optionally, the `atomic_vector` group may contain `names`, a 1-dimensional string dataset of the same dimensions as `values`.
This contains the name for each element in the atomic vector.
Names should not be missing so any placeholder attribute on `names` will not be respected.

## Height

The height of the atomic vector is defined as the extent of the sole dimension of the `atomic_vector/values` dataset in `contents.h5`.
