```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Atomic vector (", as.character(.version), ")", sep="")
```

## Overview

The `atomic_vector` format provides an on-disk representation of typed arrays, based on R's atomic vectors.

## Object metadata

The `OBJECT` file should contain an `atomic_vector` property, which is itself a JSON object with the following properties:

- `version`, a string specifying the version of the `atomic_vector` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Directory contents

The directory should contain a `contents.h5` HDF5 file.
This file should contain an `atomic_vector` group with the following attributes:

- `type`, a scalar string attribute specifying the type of the vector.
```{r, results="asis", echo=FALSE}
if (.version < package_version("1.1")) {
cat('  This should be one of `"integer"`, `"boolean"`, `"string"` or `"number"`.')
} else {
cat('  This should be one of `"integer"`, `"boolean"`, `"string"`, `"vls"` or `"number"`.')
}
```
  The datatype of the attribute should be representable by a UTF-8 encoded string.

```{r, results="asis", echo=FALSE}
if (.version < package_version("1.1")) {
cat('## Vector representation

The `atomic_vector` group should in turn contain `values`, a 1-dimensional dataset.')
} else {
cat('## Non-VLS representations

For all `type != "vls"`, the `atomic_vector` group should contain `values`, a 1-dimensional dataset.')
}
```
The HDF5 datatype of the dataset should match the `type` attribute:

- For `type = "integer"` or `"boolean"`, the datatype should be representable by a 32-bit signed integer.
  Check out the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for more details.
- For `type = "number"`, the datatype should be representable by a 64-bit IEEE754-compliant float.
  Both integer and float datatypes may be used here.
- For `type = "string"`, the datatype should be representable by a UTF-8 encoded string.

Missing values in the atomic vector are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `atomic_vector/values` dataset.
The attribute should be a scalar and should have the same datatype as `atomic_vector/values`, 
(except in the case of strings, where any datatype may be used for the attribute as long as it is compatible with a UTF-8 encoded string).
All entries in the atomic vector equal to the placeholder value should be treated as missing.
See the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for details.

For `type = "string"`, the `atomic_vector` group may optionally contain a `format` attribute.
If present, this should be a scalar attribute of a datatype that is represented by a UTF-8 encoded string.
The attribute itself should contain one of the following values:

- `"none"`: no constraints on the contents of each string.
  This is the default behaviour if no `format` is present.
- `"date"`: strings should be dates following a `YYYY-MM-DD` format.
- `"date-time"`: strings should be Internet Date/Time values following the format described in [RFC3339](https://www.rfc-editor.org/rfc/rfc3339).


```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
cat('## VLS representation
    
For `type = "vls"`, the atomic vector consists of strings stored in [**ritsuko**\'s custom variable length string (VLS) array](https://github.com/ArtifactDB/ritsuko).
Here, the `atomic_vector` group should contain the `pointers` and `heap` datasets.

- The `pointers` dataset should be a 1-dimensional dataset of a compound datatype of 2 members, `"offset"` and `"length"`.
  Each member should be of a datatype that can be represented by an unsigned 64-bit integer.
- The `heap` dataset should be a 1-dimensional dataset of unsigned 8-bit integers.

Each entry of `pointers` refers to a slice `[offset, offset + length)` of the `heap` dataset.
This slice defines a variable length UTF-8 encoded string of length `length` - unless the slice contains a null terminator, in which case the string is defined as the interval to the first null. 
Pointers may be in any order, overlapping or non-contiguous, as long as `[offset, offset + length)` lies within the boundaries of the heap.

Missing values in the atomic vector are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `atomic_vector/pointers` dataset.
The attribute should be a scalar and should be of any HDF5 string datatype that can be represented by a UTF-8 encoded string.')
}
```

## Names

Optionally, the `atomic_vector` group may contain `names`, a 1-dimensional string dataset of the same dimensions as 
```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
    cat('`atomic_vector/values` (for `type != "vls"`) or `atomic_vector/pointers` (for `type == "vls"`).')
} else {
    cat("`atomic_vector/values`.")
}
```
This contains the name for each element in the atomic vector.
The datatype should be representable by a UTF-8 encoded string.
Names should not be missing so any placeholder attribute on `names` will not be respected.

## Height

The height of the atomic vector is defined as the extent of the sole dimension of the 
```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
    cat('`atomic_vector/values` dataset (for `type != "vls"`) or `atomic_vector/pointers` dataset (for `type == "vls"`)')
} else {
    cat("`atomic_vector/values` dataset")
}
```
in `contents.h5`.
