```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Atomic vector list (", as.character(.version), ")", sep="")
```

## Overview

The `atomic_vector_list` format provides an on-disk representation of a list of atomic vectors of the same type.
This is based on the `CompressedList` subclasses for atomic vectors, as implemented in the [**IRanges** Bioconductor package](https://bioconductor.org/packages/IRanges).

## Metadata

The top-level object in the `OBJECT` file should contain an `atomic_vector_list` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `atomic_vector_list` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain a `partitions.h5` HDF5 file.
This file should contain an `atomic_vector_list` group, inside of which should be a `lengths` dataset.
The `lengths` dataset should be 1-dimensional and have a datatype that can be exactly represented by a 64-bit unsigned integer.
Each entry corresponds to an element of the atomic vector list - that is, an atomic vector -
and defines the number of consecutive elements in the `concatenated` object (see below) that can be used to construct each atomic vector.

The directory should also contain a `concatenated` subdirectory, which contains a child object of the `atomic_vector` type.
The height of this child object should be equal to the sum of lengths in the `atomic_vector_list/lengths` dataset of the `partitions.h5` file.
In effect, this child object represents the concatenation of all of the atomic vectors in the list;
reconstructing the vector for a particular list element involves taking a contiguous slice at an offset defined by the sum of the preceding `lengths`.

Optionally, the `atomic_vector_list` group may contain `names`, a 1-dimensional string dataset of the same dimensions as `values`.
This contains the name for each element of the atomic vector list.
Any HDF5 string datatype may be used here.

The directory may contain an `element_annotations` subdirectory, which contains a child object that satisfies the `DATA_FRAME` interface.
This child object should have number of rows equal to the extent of the `atomic_vector_list/lengths` dataset.
Each row corresponds to a list element and contains additional annotations for that element.

The directory may contain an `other_annotations` subdirectory, which contains a child object that satisfies the `SIMPLE_LIST` interface.
This holds extra annotations for the entire list.

## Height

The height of the atomic vector list is defined as the extent of the sole dimension of the `atomic_vector_list/lengths` dataset in `partitions.h5`.
