```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# BAM file (", as.character(.version), ")", sep="")
```

## Overview

The `bam_file` format specifies the expected representation for a BAM file.

## Object metadata

The `OBJECT` file should contain a `bam_file` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `bam_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Directory structure

The directory should contain `file.bam`, the BAM file.
The file should be formatted as described in the [BAM specification](https://samtools.github.io/hts-specs/SAMv1.pdf).

The directory may optionally contain `file.bam.bai` and/or `file.bam.csi`, which should be indices into `file.bam`.
`file.bam.bai` should follow the format described in the [BAM specification](https://samtools.github.io/hts-specs/SAMv1.pdf),
while `file.bam.csi` should follow the [CSI specification](https://samtools.github.io/hts-specs/CSIv1.pdf).
Both files may be present in a single directory, though it is generally only necessary to provide one.
