```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# BCF file (", as.character(.version), ")", sep="")
```

## Overview

The `bcf_file` format specifies the expected directory organization for BCF files.

## Metadata

The top-level object in the `OBJECT` file should contain a `bcf_file` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `bcf_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain `file.bcf`, the BCF file.
The file should be formatted as described in the BCF specification - 
any version is supported, but it is generally recommended to use version 2 or higher (https://samtools.github.io/hts-specs/BCFv2_qref.pdf).

The directory may optionally contain `file.bcf.tbi` and/or `file.bcf.csi`, which should be indices into `file.bcf`.
`file.bcf.tbi` should follow the format described in the [Tabix specification](https://samtools.github.io/hts-specs/tabix.pdf),
while `file.bcf.csi` should follow the [CSI specification](https://samtools.github.io/hts-specs/CSIv1.pdf).
Both files may be present in a single directory, though it is generally only necessary to provide one.
