```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# BED file (", as.character(.version), ")", sep="")
```

## Overview

The `bed_file` format specifies the expected directory organization for BED files.

## Metadata

The top-level object in the `OBJECT` file should contain a `bed_file` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `bed_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- `indexed`, a boolean indicating whether the BED file has been indexed.

## Files

If `indexed = false`, the directory should contain `file.bed.gz`, a Gzip-compressed BED file.
The file should be formatted as described in the [BED specification](https://samtools.github.io/hts-specs/BEDv1.pdf).

If `indexed = true`, the directory should contain `file.bed.bgz`, a BGZF-compressed file following the BED format.
The directory should also contain `file.bed.bgz.tbi`, a [Tabix file ](https://samtools.github.io/hts-specs/tabix.pdf) that indexes into `file.bed.bgz`.
