```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Compressed sparse matrix (", as.character(.version), ")", sep="")
```

## Overview

The `compressed_sparse_matrix` format provides an on-disk representation of a sparse matrix in the compressed sparse row/column format.

## Metadata

The top-level object in the `OBJECT` file should contain an `compressed_sparse_matrix` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `compressed_sparse_matrix` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain a `matrix.h5` HDF5 file, which in turn should contain an `compressed_sparse_matrix` group.
The group should contain `shape`, a 1-dimensional dataset of any type that can be exactly represented by a 64-bit unsigned integer.
`compressed_sparse_matrix/shape` should have length 2 where the entries specify the number of rows and columns, respectively, of the matrix.

The `compressed_sparse_matrix` group should have a `layout` attribute.
This should be a scalar string dataset that is either `"CSR"` or `"CSC"`, specifying whether the matrix is compressed sparse row (CSR) or column (CSC).
Any HDF5 string datatype may be used here.

The `compressed_sparse_matrix` group should contain `data`, a 1-dimensional dataset containing the values for the structural non-zeroes.
This should have a `type` attribute specifying the type of the sparse matrix and the expected HDF5 datatype of the dataset itself:

- For `type = "integer"` or `"boolean"`, the datatype should be any integer type that can be exactly represented by a 32-bit signed integer.
  Check out the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for more details.
- For `type = "number"`, the datatype should be any integer/float type that can be exactly represented by a 64-bit float.

Missing values in the matrix are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `compressed_sparse_matrix/data` dataset.
The attribute should be scalar and should have the same datatype as `compressed_sparse_matrix/data`.
All values in the dataset equal to the placeholder should be treated as missing.

The `compressed_sparse_matrix` group should contain `indices`, a 1-dimensional dataset containing the row indices (for CSC) or column indices (for CSR) for the structural non-zeroes.
This should have a datatype that can be exactly represented by a 64-bit unsigned integer.
The length of `compressed_sparse_matrix/indices` should be equal to that of `compressed_sparse_matrix/data`, as each entry of the former contains the row/column index of the latter.
All values should be less than the number of rows (for CSC) or columns (for CSR) in `compressed_sparse_matrix/shape`.

The `compressed_sparse_matrix` group should contain `indptr`, a 1-dimensional dataset containing the pointer offsets for the columns (for CSC) or rows (for CSR).
The length of this dataset should be equal to the number of columns plus 1 (for CSC) or the number of rows plus 1 (for CSR).
This should have a datatype that can be exactly represented by a 64-bit unsigned integer.
The first value should be zero, the last value should be equal to the length of `compressed_sparse_matrix/data`, and all values should be sorted in non-decreasing order.

Given an index `i` (for rows in CSR, columns in CSC), `indptr[i]` contains the position of the first structural non-zero for that row/column in `compressed_sparse_matrix/data`.
Similarly, `indptr[i+1]` contains the one-past-the-end position for that row/column, such that the range between `indptr[i]` and `indptr[i+1]` contains all structural non-zeros in the row/column.
Given a particular range, all values in `compressed_sparse_matrix/indices` inside that range should be sorted in strictly increasing order.

Optionally, the `compressed_sparse_matrix` group may contain a `names` subgroup.
This may contain any number of 1-dimensional string datasets named after the corresponding dimension, e.g., `0` for rows, `1` for columns.
Each dataset contains the names for its dimension and should have a length equal to the extent of that dimension.
Any HDF5 string datatype may be used in each dataset.

## Height

The height of the compressed sparse matrix is defined as the number of rows, as specified in `compressed_sparse_matrix/shape`.

## Dimensions

The dimensions of the compressed sparse matrix are taken from `compressed_sparse_matrix/shape`.
