```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Data frame (", as.character(.version), ")", sep="")
```

## Overview

The `data_frame` format provides an on-disk representation of a data frame, based on the `DataFrame` class from the [**S4Vectors** Bioconductor package](https://bioconductor.org/packages/S4Vectors).

## Metadata

The top-level object in the `OBJECT` file should contain a `data_frame` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `data_frame` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain a `basic_contents.h5` HDF5 file, which in turn should contain an `data_frame` group.
The group should have a `row-count` attribute, which should be a scalar dataset of any datatype that can be exactly represented by a 64-bit unsigned integer.
The value of this attribute specifies the number of rows in the data frame.

The `data_frame` group should contain `column_names`, a 1-dimensional string dataset containing the column names of the data frame.
Any HDF5 string datatype may be used here.
Column names should not be empty or duplicated.
The length of this dataset defines the number of columns of the data frame.

The `data_frame` group may optionally contain `row_names`, a 1-dimensional string dataset containing the row names of the data frame if any are present.
Any HDF5 string datatype may be used here.
The length of this dataset should be equal to the number of rows.

The `data_frame` group should contain the `data` subgroup, which stores all "basic" columns.
Each basic column is represented by a HDF5 group or dataset named after the positional index of the column, e.g., `"0"` for the first column, `"1"` for the second column, and so on.
See below on the expected representation of each basic column.

The directory may also contain a `other_contents` subdirectory, which stores all "non-basic" columns.
Each non-basic column is represented by a subdirectory within `other_contents` that is named after the positional index of the column, e.g., `"2"` for the third column, `"3"` for the fourth column, and so on.
See below on the expected representation of each non-basic column.

Each column in the data frame should be represented exactly once across `basic_contents.h5` and `other_contents`.

The directory may contain an `element_annotations` subdirectory, which contains a child object that satisfies the `DATA_FRAME` interface.
This child object should have number of rows equal to the number of columns, i.e., the length of the `data_frame/column_names` dataset in `basic_contents.h5`.
Each row of this child object corresponds to a column of the data frame and contains additional annotations for that column.

The directory may contain an `other_annotations` subdirectory, which contains a child object that satisfies the `SIMPLE_LIST` interface.
This holds extra annotations for the entire data frame.

## Columns

### Basic, not factor

We now consider the basic column X, i.e., the X-th column in the data frame.
If this column is not a factor, it is represented by a HDF5 dataset at `data_frame/data/X`.
This dataset should be a 1-dimensional and of length equal to the number of rows.
It should have a `type` attribute containing a scalar string - any HDF5 string datatype may be used here.
The value of the attribute specifies the type of the column and the expected HDF5 datatype of the dataset itself:

- For `type = "integer"` or `"boolean"`, the datatype should be any integer type that can be exactly represented by a 32-bit signed integer.
  Check out the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for more details.
- For `type = "number"`, the datatype should be any integer/float type that can be exactly represented by a 64-bit float.
- For `type = "string"`, the datatype should be any string type that can be represented by a UTF-8 encoded string.

For `type = "string"`, the `data_frame/data/X` dataset may optionally contain a `format` attribute.
This should be set to:

- `"none"`: no constraints on the contents of each string.
  This is the default behaviour if no `format` is present.
- `"date"`: strings should be dates following a `YYYY-MM-DD` format.
- `"date-time"`: strings should be Internet Date/Time values following the format described in [RFC3339](https://www.rfc-editor.org/rfc/rfc3339).

Missing values are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `data_frame/data/X` dataset.
The attribute should be scalar and have the same datatype as `data_frame/data/X`, except in the case of strings where any string datatype may be used.
All values in the dataset equal to the placeholder should be treated as missing.

### Factor

If column X is a factor, it is instead represented by a HDF5 group at `data_frame/data/X`.
This group should have a `type` attribute containing a scalar string - any HDF5 string datatype may be used here.
`type` should be equal to `"factor"`.

The `data_frame/data/X` group should contain `levels`, a 1-dimensional string dataset containing the factor levels.
Any HDF5 string datatype may be used here.
All levels should be unique.

The `data_frame/data/X` group should contain `codes`, a 1-dimensional dataset containing the factor codes. 
The length of this dataset should be equal to the number of rows, and the datatype should be exactly represented by a 64-bit unsigned integer.
Values of this dataset should either be less than the number of levels or equal to the placeholder value.

Missing factor entries are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `data_frame/data/X/codes` dataset.
The attribute should be scalar and have the same datatype as `data_frame/data/X/codes`.
All values in the dataset equal to the placeholder should be treated as missing.

This group may also have an `ordered` attribute, which should be a scalar of any datatype that fits into a 32-bit signed integer.
If present, a non-zero value indicates that the factor levels should be treated as ordered.
Otherwise, the levels are treated as unordered.

### Other

We consider the non-basic column Y, i.e., the X-th column in the data frame.
This may be any object type that is supported by **takane** and has a concept of "height";
a common use case is that of nested data frames where one data frame is a column of another data frame.
The subdirectory at `other_contents/Y` holds the on-disk representation of this non-basic column, so `other_contents/Y` is now a child object of the enclosing data frame.
The height of the child object should be equal to the number of rows in the data frame, as defined from the `row-count` attribute of the `data_frame` group inside `basic_contents.h5`.

## Height

The height of the data frame is defined as the number of rows, as specified in `row-count`.

## Dimensions

The dimensions of the data frame is defined as the number of rows (`row-count`) and columns (the length of `column_names`).

## Interfaces

The `data_frame` object satisfies the `DATA_FRAME` interface.
