```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Dense array (", as.character(.version), ")", sep="")
```

## Overview

The `dense_array` format provides an on-disk representation of a dense array.

## Object metadata

The `OBJECT` file should contain a `dense_array` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `dense_array` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Directory contents

The directory should contain an `array.h5` HDF5 file.
This file should contain a `dense_array` group with the following attributes:

- `type`, a scalar string attribute specifying the type of the array 
```{r, results="asis", echo=FALSE}
if (.version < package_version("1.1")) {
cat('  This should be one of `"integer"`, `"boolean"`, `"string"` or `"number"`.')
} else {
cat('  This should be one of `"integer"`, `"boolean"`, `"string"`, `"vls"` or `"number"`.')
}
```
  The datatype of the attribute should be representable by a UTF-8 encoded string.

```{r, results="asis", echo=FALSE}
if (.version < package_version("1.1")) {
cat('## Vector representation

The `dense_array` group should contain `values`, a 1-dimensional dataset.')
} else {
cat('## Non-VLS representations

For all `type != "vls"`, the `dense_array` group should contain `values`, a 1-dimensional dataset.')
}
```
The group should contain `data`, an N-dimensional dataset containing the values of the dense array.
The datatype of this dataset depends on `type`:

- For `type = "integer"` or `"boolean"`, the datatype should be representable by a 32-bit signed integer.
  Check out the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for more details.
- For `type = "number"`, the datatype should be representable by a 64-bit IEEE754-compliant float.
  Both integer and float datatypes may be used here.
- For `type = "string"`, the datatype should be representable by a UTF-8 encoded string.

Missing values in the array are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `dense_array/data` dataset.
The attribute should be a scalar and should have the same datatype as `dense_array/data`, 
(except in the case of strings, where any datatype may be used for the attribute as long as it is compatible with a UTF-8 encoded string).
All values in the dataset equal to the placeholder should be treated as missing.
See the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for details.

```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
cat('## VLS representation
    
For `type = "vls"`, the dense array consists of strings stored in [**ritsuko**\'s custom variable length string (VLS) array](https://github.com/ArtifactDB/ritsuko).
Here, the `dense_array` group should contain the `pointers` and `heap` datasets.

- The `pointers` dataset should be an N-dimensional dataset of a compound datatype of 2 members, `"offset"` and `"length"`.
  Each member should be of a datatype that can be represented by an unsigned 64-bit integer.
- The `heap` dataset should be a 1-dimensional dataset of unsigned 8-bit integers.

Each entry of `pointers` refers to a slice `[offset, offset + length)` of the `heap` dataset.
This slice defines a variable length UTF-8 encoded string of length `length` - unless the slice contains a null terminator, in which case the string is defined as the interval to the first null. 
Pointers may be in any order, overlapping or non-contiguous, as long as `[offset, offset + length)` lies within the boundaries of the heap.

Missing values in the dense array are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `dense_array/pointers` dataset.
The attribute should be a scalar and should be of any HDF5 string datatype that can be represented by a UTF-8 encoded string.')
}
```

## Dimension names

Optionally, the `dense_array` group may contain a `names` subgroup.
This may contain any number of 1-dimensional string datasets named after the corresponding dimension of `dense_array/data`, e.g., `0` for the first dimension, `1` for the second dimension, and so on.
Each dataset contains the names for its dimension and should have a length equal to the extent of that dimension.
The datatype of each dataset should be representable by a UTF-8 encoded string.

## Transposition

Optionally, the `dense_array` group may have a `transposed` attribute.
This should be scalar and have a datatype that is representable by a 32-bit signed integer.
Any non-zero value indicates that the dense array is stored in a transposed format,
i.e., the first HDF5 dimension is actually the last dimension in the original array.
This is necessary for efficiently reading/writing from frameworks that use, e.g., column-major layouts to represent matrices, as opposed to HDF5's row-major layout.

## Height

The height of the dense array is defined as the extent of the first dimension of the
```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
    cat('`dense_array/values` dataset (for `type != "vls"`) or `dense_array/pointers` dataset (for `type == "vls"`)')
} else {
    cat("`dense_array/values` dataset")
}
```
in `array.h5`.
If the array is transposed, the extent of the last dimension is used instead.

## Dimensions

The dimensions of the dense array are defined as the dimensions of the 
```{r, results="asis", echo=FALSE}
if (.version >= package_version("1.1")) {
    cat('`dense_array/values` dataset (for `type != "vls"`) or `dense_array/pointers` dataset (for `type == "vls"`)')
} else {
    cat("`dense_array/values` dataset")
}
```
in `array.h5`.
If the array is transposed, the dimensions are reversed.
