```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# FASTA file (", as.character(.version), ")", sep="")
```

## Overview

The `fasta_file` format specifies the expected directory organization for FASTA files.

## Metadata

The top-level object in the `OBJECT` file should contain a `fasta_file` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `fasta_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- `indexed`, a boolean indicating whether the FASTA file has been indexed.

## Files

If `indexed = false`, the directory should contain `file.fasta.gz`, a Gzip-compressed FASTA file.

If `indexed = true`, the directory should contain `file.fasta.bgz`, a BGZF-compressed FASTA file.
The directory should also contain `file.fasta.fai`, the FASTA index file containing offsets into the uncompressed `file.fasta`;
and `file.fasta.bgz.gzi`, the BGZF index file containing the byte offset mappings the uncompressed file to the compressed file.
