```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Genomic ranges list (", as.character(.version), ")", sep="")
```

## Overview

The `genomic_ranges_list` format provides an on-disk representation of a list of genomic ranges of the same type.
This is based on the `GRangesList` from the [**GenomicRanges** Bioconductor package](https://bioconductor.org/packages/GenomicRanges).

## Metadata

The `OBJECT` file should contain a `genomic_ranges_list` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `genomic_ranges_list` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Directory structure

The directory should contain a `partitions.h5` HDF5 file.
This file should contain an `genomic_ranges_list` group, inside of which should be a `lengths` dataset.
The `lengths` dataset should be 1-dimensional and have a datatype that can be exactly represented by a 64-bit unsigned integer.
Each entry corresponds to an element of the genomic ranges list - that is, a genomic ranges object -
and defines the number of consecutive elements in the `concatenated` object (see below) that can be used to construct each genomic ranges object.

The directory should also contain a `concatenated` subdirectory, which contains a child object of the [`genomic_ranges`](../genomic_ranges) type.
The height of this child object should be equal to the sum of lengths in the `genomic_ranges_list/lengths` dataset of the `partitions.h5` file.
In effect, this child object represents the concatenation of all of the genomic ranges in the list;
reconstructing the vector for a particular list element involves taking a contiguous slice at an offset defined by the sum of the preceding `lengths`.

Optionally, the `genomic_ranges_list` group may contain `names`, a 1-dimensional string dataset of the same dimensions as `genomic_ranges_list/lengths`.
This contains the name for each element of the genomic ranges list.
Any HDF5 string datatype may be used here.

The directory may contain an `element_annotations` subdirectory, which contains a child object that satisfies the `DATA_FRAME` interface.
This child object should have number of rows equal to the extent of the `genomic_ranges_list/lengths` dataset.
Each row corresponds to a list element and contains additional annotations for that element.

The directory may contain an `other_annotations` subdirectory, which contains a child object that satisfies the `SIMPLE_LIST` interface.
This holds extra annotations for the entire list.

## Height

The height of the genomic ranges list is defined as the extent of the sole dimension of the `genomic_ranges_list/lengths` dataset in `partitions.h5`.
