```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Image file (", as.character(.version), ")", sep="")
```

## Overview

The `image_file` format specifies the expected representation for an image file in one of the standard formats.

## Object metadata

The `OBJECT` file should contain a `bam_file` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `bam_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- `format`, a string specifying the format of the file.
  This may be one of `"PNG"`, `"TIFF"`, `"JPEG"`, `"GIF"` or `"WEBP"`.

## Directory structure

The directory should contain:

- `file.png`, a PNG file if `format` is `"PNG"`.
- `file.tiff`, a TIFF file if `format` is `"TIFF"`.
- `file.jpg`, a JPEG file if `format` is `"JPEG"`.
- `file.gif`, a GIF file if `format` is `"GIF"`.
- `file.webp`, a WEBP file if `format` is `"WEBP"`.

All formats are validated based on their magic numbers.
