```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Multi-sample dataset (", as.character(.version), ")", sep="")
```

## Overview

The `multi_sample_dataset` format provides an on-disk representation of a multi-sample dataset.
It is based on the `MultiAssayExperiment` class from the [**MultiAssayExperiment** bioconductor package](https://bioconductor.org/packages/MultiAssayExperiment).

## Metadata

The top-level object in the `OBJECT` file should contain a `multi_sample_dataset` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `multi_sample_dataset` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain a `sample_data` subdirectory, which contains a child object that satisfies the `DATA_FRAME` interface.
The number of samples in the multi-sample dataset is determined as the number of rows in the `sample_data` data frame.

The directory may contain an `experiments` subdirectory.
The subdirectory should contain a `names.json` file, which contains a JSON-formatted array of strings.
Each string is the name of an experiment and should be non-empty and unique.
The `experiments` subdirectory should contain one subdirectory for each experiment, named after the positional index in the JSON array, e.g., `"0"` for the first experiment, `"1"` for the second experiment, and so on.
Each subdirectory of `experiments` should be a child object that satisfies the `SUMMARIZED_EXPERIMENT` interface.
If no `experiments` subdirectory is present, it is assumed that the multi-sample dataset contains no experiments.

If the directory contains an `experiments` subdirectory, it should further contain a `sample_map.h5` HDF5 file.
This should contain a `multi_sample_dataset` group, which further contains one 1-dimensional integer dataset per experiment in `experiments`.
Each dataset is named after the positional index of the experiment (e.g., `"0"`, `"1"`, as described above) and should have length equal to the number of columns of its experiment.
The datatype should be exactly representable by a 64-bit unsigned integer.
The values in each dataset are indices into the rows of the `sample_data` data frame, and thus should be less than the number of samples.

The directory may contain an `other_data` subdirectory, which contains a child object that satisfies the `SIMPLE_LIST` interface.
This holds extra annotations for the entire multi-sample dataset.
