```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# RDS file (", as.character(.version), ")", sep="")
```

## Overview

The `rds_file` format specifies the expected representation for a RDS file.
This is the serialization format for the R programming language and is best avoided for applications that need interoperability between languages;
the format itself requires some guesswork to parse, and there are no checks on the content of the file (e.g., users can save any arbitrary objects).
Nonetheless, we list it here as it may be occasionally necessary to save objects that have no better representation in **takane**.

## Object metadata

The `OBJECT` file should contain a `rds_file` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `rds_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Directory structure

The directory should contain `file.rds`, a Gzip-compressed RDS file in XDR binary format.
The file should be formatted as described in the [R internals documentation](https://cran.r-project.org/doc/manuals/r-release/R-ints.html#Serialization-Formats).
