```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Sequence information (", as.character(.version), ")", sep="")
```

## Overview

The `sequence_information` format provides an on-disk representation of a data frame of reference sequence information. 
It is based on the `Seqinfo` class from the [**GenomeInfoDb** Bioconductor package](https://bioconductor.org/packages/GenomeInfoDb).

## Metadata

The `OBJECT` file should contain a `sequence_information` property, itself a JSON object with the following properties:

- `version`, a string specifying the version of the `sequence_information` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

## Files

The directory should contain an `info.h5` HDF5 file, which in turn should contain a `sequence_information` group.
This group should in turn contain several 1-dimensional datasets, all of which should be of the same length.
The length of each dataset represents the number of reference sequences
The following datasets should be present:

- `name`, a dataset of any HDF5 string datatype, containing the names of the reference sequences.
  All values should be unique.
- `length`, a dataset of any datatype that can be exactly represented by a 64-bit unsigned integer.
  Each value contains the length of each sequence in base pairs.
- `circular`, a dataset of any datatype that can be exactly represented by a 32-bit signed integer.
  A non-zero value indicates that the corresponding reference sequence is circular.
- `genome`, a dataset of any HDF5 string datatype, containing the identifier of the genome that contains the sequence.
  This may contain duplicate values across sequences.

`length`, `circular` and `genome` may contain missing values, which are represented by placeholder values defined in the `missing-value-placeholder` attribute of each dataset.
The attribute should be a scalar and should have the same datatype as its dataset, except in the case of strings where any string datatype may be used.
All values in the dataset equal to the placeholder should be treated as missing.
See the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for details.
