```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Simple list (", as.character(.version), ")", sep="")
```

## Overview

The `simple_list` format specifies the on-disk representation of a simple list.
This is based on an ordinary R list.

## Metadata

The top-level object in the `OBJECT` file should contain a `simple_list` property.
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `fasta_file` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- (optional) `format`, a string specifying the format to use.
  This should be one of `"hdf5"` or `"json.gz"`.
  If not provided, it is assumed to be `"hdf5"`.

## Files

If `format = "hdf5"`, the directory should contain a `list_contents.h5` file.
This is a HDF5 file that follows the [**uzuki2**](https://github.com/ArtifactDB/uzuki2) HDF5 specification for R lists.

If `format = "hdf5"`, the directory should contain a `list_contents.json.gz` file.
This is a Gzip-compressed JSON file that follows the [**uzuki2**](https://github.com/ArtifactDB/uzuki2) JSON specification for R lists.

The directory may contain an `other_contents` subdirectory, containing external list items as further subdirectories.
Each subdirectory in `other_contents` represents a child object and should be named after its external list index, which are consecutive zero-based indices, i.e., `"0"`, `"1"`, and so on.
Every external list index should be used at least once by the list representation in `list_contents.h5` or `list_contents.json.gz`.

## Height

The height of the sequence string set is defined as the length of the list in `list_contents.h5` or `list_contents.json.gz`.
Only the length of the top-level list is considered here.
