```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Single cell experiment (", as.character(.version), ")", sep="")
```

## Overview

The `single_cell_experiment` format provides an on-disk representation of a single cell experiment. 
It is based on the `SingleCellExperiment` class from the [**SingleCellExperiment** bioconductor package](https://bioconductor.org/packages/SingleCellExperiment).

## Metadata

The `OBJECT` file should contain a `single_cell_experiment` property, which is itself a JSON object with the following properties:

- `version`, a string specifying the version of the `single_cell_experiment` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- (optional) `main_experiment_name`, a string containing the name of the main experiment.
  If provided, this should not overlap with any names of the alternative experiments in the `alternative_experiments` subdirectory.

In addition, the `OBJECT` file should contain all properties expected for the [`ranged_summarized_experiment`](../ranged_summarized_experiment) format, i.e., a `ranged_summarized_experiment` property.
This transitively includes the properties that are expected of the [`summarized_experiment`](../summarized_experiment) format. 

## Files

The directory should contain all files expected of a [`ranged_summarized_experiment`](../ranged_summarized_experiment) format.
This includes the transitive expectations of files that should be present in a [`summarized_experiment`](../summarized_experiment) format.
The dimensions of the single cell experiment are determined from the `summarized_experiment`-based files.

The directory may contain a `reduced_dimensions` subdirectory.
The subdirectory should contain a `names.json` file, which contains a JSON-formatted array of strings.
Each string is the name of a dimensionality reduction result and should be non-empty and unique.
The `reduced_dimensions` subdirectory should contain one subdirectory for each dimensionality reduction result,
named after the result's positional index in the JSON array, e.g., `0` for the first result, `1` for the second result, and so on.
Each subdirectory of `reduced_dimensions` should be a child object that has some concept of dimensions,
where the extent of the first dimension is equal to the number of columns of the single cell experiment.
If no `reduced_dimensions` subdirectory is present, it is assumed that the single cell experiment contains no reduced dimensions.

The directory may contain an `alternative_experiments` subdirectory.
The subdirectory should contain a `names.json` file, which contains a JSON-formatted array of strings.
Each string is the name of an alternative experiment and should be non-empty and unique.
The `alternative_experiments` subdirectory should contain one subdirectory for each alternative experiment,
named after the experiment's positional index in the JSON array, e.g., `0` for the first experiment, `1` for the second experiment, and so on.
Each subdirectory of `reduced_dimensions` should be a child object that satisfies the `SUMMARIZED_EXPERIMENT` interface,
with number of columns equal to the number of columns of the single cell experiment.
If no `alternative_experiments` subdirectory is present, it is assumed that the single cell experiment contains no alternative experiments.

Note that any names in the `reduced_dimensions` or `alternative_experiments` should not be used as the column names of the ranged summarized experiment, 
as this is determined from the `summarized_experiment`-based files instead.

## Height

The height of the `single_cell_experiment` is defined in the same manner as that of the underlying [`summarized_experiment`](../summarized_experiment).

## Dimensions

The dimensions of the `single_cell_experiment` is defined in the same manner as that of the [`summarized_experiment`](../summarized_experiment).

## Interfaces

The `single_cell_experiment` satisfies the `SUMMARIZED_EXPERIMENT` interface.
