```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# String factor (", as.character(.version), ")", sep="")
```

## Overview

The `string_factor` format provides an on-disk representation of a factor (i.e., enumerated type).
This is based on R's factor.

## Metadata

The `OBJECT` file should contain a `string_factor` property, which is itself a JSON object with the following properties:

- `version`, a string specifying the version of the `string_factor` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```

### Files

The directory should contain the `contents.h5` HDF5 file.
This should contain a `sting_factor` group, which should in turn contain `levels`.
This is a 1-dimensional string dataset containing the factor levels.
The datatype should be representable by a UTF-8 encoded string.
All levels should be unique.

The `string_factor` group should also contain `codes`, a 1-dimensional dataset containing the 0-indexed factor codes. 
The length of this dataset should be equal to the number of rows, and the datatype should be representable by a 64-bit unsigned integer.
Values of this dataset should either be less than the number of levels or equal to the placeholder value.

Missing factor entries are represented by placeholder values, defined in the `missing-value-placeholder` attribute of the `string_factor/codes` dataset.
The attribute should be scalar and have the same datatype as `string_factor/codes`.
All values in the dataset equal to the placeholder should be treated as missing.
See the [HDF5 policy draft (v0.1.0)](https://github.com/ArtifactDB/Bioc-HDF5-policy/tree/v0.1.0) for details.

The `string_factor` group may also have an `ordered` attribute, which should be a scalar of any datatype that fits into a 32-bit signed integer.
If present, a non-zero value indicates that the factor levels should be treated as ordered.
Otherwise, the levels are treated as unordered.

Optionally, the `string_factor` group may contain `names`, a 1-dimensional string dataset of the same dimensions as `codes`.
This contains the name for each element in the factor.
The datatype should be representable by a UTF-8 encoded string.
Names should not be missing so any placeholder attribute on `names` will not be respected.

## Height

The height of the data frame is defined as the extent of the sole dimension of the `string_factor/codes` dataset in `contents.h5`.
