```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Summarized experiment (", as.character(.version), ")", sep="")
```

## Overview

The `summarized_experiment` format provides an on-disk representation of a summarized experiment. 
It is based on the `SummarizedExperiment` class from the [**SummarizedExperiment** bioconductor package](https://bioconductor.org/packages/SummarizedExperiment).

## Object metadata

The `OBJECT` file should contain an `summarized_experiment` property, which is itself a JSON object with the following properties:

- `version`, a string specifying the version of the `summarized_experiment` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- `dimensions`, an array of the dimensions of the summarized experiment.
  This should contain two non-negative integers specifying the number of rows and columns, respectively.

## Directory structure

The directory may contain an `assays` subdirectory.
The subdirectory should contain a `names.json` file, which contains a JSON-formatted array of strings.
Each string is the name of an assay and should be non-empty and unique.
The `assays` subdirectory should contain one subdirectory for each assay,
named after the assay's positional index in the JSON array, e.g., `0` for the first assay, `1` for the second assay, and so on.
Each subdirectory of `assays` should be a child object that has some concept of "dimensions",
where the number of dimensions is at least 2 and the first two dimensions are equal to those of the summarized experiment.
If no `assays` subdirectory is present, it is assumed that no assays are available.

The directory may contain a `row_data` subdirectory, containing additional annotations for the rows.
This should contain a child object that satisfies the `DATA_FRAME` contract.
The number of rows in this object should be equal to the number of rows in the summarized experiment.
If this object contains row names, these should be treated as the row names of the summarized experiment.
If no `row_data` subdirectory is present, it can be assumed that no additional row annotations are available.

The directory may contain a `column_data` subdirectory, containing additional annotations for the columns. 
This should contain a child object that satisfies the `DATA_FRAME` contract.
The number of rows in this object should be equal to the number of columns in the summarized experiment.
If this object contains row names, these should be treated as the column names of the summarized experiment.
If no `column_data` subdirectory is present, it can be assumed that no additional column annotations are available.

The directory may contain an `other_data` subdirectory, which contains a child object that satisfies the `SIMPLE_LIST` interface.
This holds extra annotations for the entire summarized experiment. 

## Height

The height of the summarized experiment is defined as the number of rows, determined from the `dimensions` property in the object metadata.

## Dimensions

The dimensions of the summarized experiment is taken from the `dimensions` property in the object metadata.

## Interfaces

The summarized experiment satisfies the `SUMMARIZED_EXPERIMENT` interface.
