```{r, results="hide", echo=FALSE}
knitr::opts_chunk$set(error=FALSE)
if (!exists(".version")) {
    .version <- package_version("1.0")
}
```

```{r, results="asis", echo=FALSE}
cat("# Summarized experiment (", as.character(.version), ")", sep="")
```

## Overview

The `vcf_experiment` format provides an on-disk representation of a VCF-derived summarized experiment. 
It is based on the `VCF` subclasses from the [**VariantAnnotation** bioconductor package](https://bioconductor.org/packages/VariantAnnotation).

## Metadata

The top-level object in the `OBJECT` file should contain a `vcf_experiment` property, 
This is a JSON object with the following properties:

- `version`, a string specifying the version of the `vcf_experiment` format.
```{r, results="asis", echo=FALSE}
cat("  This should be set to \"", as.character(.version), "\".", sep="")
```
- `dimensions`, an array of the dimensions of the VCF experiment.
  This should contain two non-negative integers specifying the number of rows and columns, respectively.
- (optional) `expanded`, a boolean indicating whether loci with multiple alternative alleles are expanded into multiple records in the VCF file.
  If missing, this defaults to `false`.

## Files

The directory should contain `file.vcf.gz`, a Gzip-compressed VCF file.
The number of records should be equal to the number of rows in the `dimensions` metadata field.
The number of samples should be equal to the number of columns in the `dimensions` metadata field.
If `expanded = true`, each record should contain no more than one alternative allele.
The VCF file should be newline-terminated.

## Height

The height of the summarized experiment is defined as the number of rows, determined from the `dimensions` property in the object metadata.

## Dimensions

The dimensions of the summarized experiment is taken from the `dimensions` property in the object metadata.
